---
title: "Assignment 2 - Language Development in ASD - Part 1 - Explaining development"
author: "[YOUR NAME]"
date: "[DATE]"
output: html_document
---
    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(include = FALSE)
```

# Assignment 2

In this assignment you will have to discuss a few important questions (given the data you have). More details below. The assignment submitted to the teachers consists of:
- a report answering and discussing the questions (so we can assess your conceptual understanding and ability to explain and critically reflect)
- a link to a git repository with all the code (so we can assess your code)

Part 1 - Basic description of language development
- Describe your sample (n, age, gender, clinical and cognitive features of the two groups) and critically assess whether the groups (ASD and TD) are balanced
- Describe linguistic development (in terms of MLU over time) in TD and ASD children (as a function of group). 
- Describe how parental use of language (in terms of MLU) changes over time. What do you think is going on?
- Include individual differences in your model of language development (in children). Identify the best model.

Part 2 - Model comparison
- Discuss the differences in performance of your model in training and testing data
- Which individual differences should be included in a model that maximizes your ability to explain/predict new data?
- Predict a new kid's performance (Bernie) and discuss it against expected performance of the two groups

Part 3 - Simulations to plan a new study
- Report and discuss a power analyses identifying how many new kids you would need to replicate the results

The following involves only Part 1.

## Learning objectives

- Summarize and report data and models
- Critically apply mixed effects (or multilevel) models
- Explore the issues involved in feature selection


# Quick recap
Autism Spectrum Disorder is often related to language impairment. However, this phenomenon has not been empirically traced in detail:
i) relying on actual naturalistic language production,  ii) over extended periods of time.

We therefore videotaped circa 30 kids with ASD and circa 30 comparison kids (matched by linguistic performance at visit 1) for ca. 30 minutes of naturalistic interactions with a parent. We repeated the data collection 6 times per kid, with 4 months between each visit. We transcribed the data and counted: 
i) the amount of words that each kid uses in each video. Same for the parent.
ii) the amount of unique words that each kid uses in each video. Same for the parent.
iii) the amount of morphemes per utterance (Mean Length of Utterance) displayed by each child in each video. Same for the parent. 

This data is in the file you prepared in the previous class. 

NB. A few children have been excluded from your datasets. We will be using them next week to evaluate how good your models are in assessing the linguistic development in new participants.

This RMarkdown file includes 
1) questions (see above). Questions have to be answered/discussed in a separate document that you have to directly submit on Blackboard.
2) A break down of the questions into a guided template full of hints for writing the code to solve the exercises. Fill in the code and the paragraphs as required. Then report your results in the doc for the teachers.

REMEMBER that you will have to have a github repository for the code and submit the answers to Blackboard without code (but a link to your github/gitlab repository). This way we can check your code, but you are also forced to figure out how to report your analyses :-)

Before we get going, here is a reminder of the issues you will have to discuss in your report:

1- Describe your sample (n, age, gender, clinical and cognitive features of the two groups) and critically assess whether the groups (ASD and TD) are balanced
2- Describe linguistic development (in terms of MLU over time) in TD and ASD children (as a function of group). 
3- Describe how parental use of language (in terms of MLU) changes over time. What do you think is going on?
4- Include individual differences in your model of language development (in children). Identify the best model.

# Let's go

### Loading the relevant libraries

Load necessary libraries : what will you need?
- e.g. something to deal with the data
- e.g. mixed effects models
- e.g. something to plot with

```{r Load Libraries, include = FALSE}

pacman::p_load(tidyverse,janitor, lme4, ggplot2)

```

### Define your working directory and load the data
If you created a project for this class and opened this Rmd file from within that project, your working directory is your project directory.

If you opened this Rmd file outside of a project, you will need some code to find the data:
- Create a new variable called locpath (localpath)
- Set it to be equal to your working directory
- Move to that directory (setwd(locpath))
- Load the data you saved last time (use read_csv(fileName))

```{r Load Data, include = FALSE}

df <- read_csv("/Users/al/RStudio/Semester 3/Experimental-Methods-3/Assignment 2/methods3_A2_P1/merged_df_assignment1.csv")

df$X1 <- NULL

```

### Characterize the participants (Exercise 1)

Identify relevant variables: participants demographic characteristics, diagnosis, ADOS, Verbal IQ, Non Verbal IQ, Socialization, Visit, Number of words used, Number of unique words used, mean length of utterance in both child and parents.

Make sure the variables are in the right format.

Describe the characteristics of the two groups of participants and whether the two groups are well matched.

```{r descriptive stats, include = FALSE}

# the number, age, gender, clinical and cognitive features of children in the groups ASD and TD.
df <- group_by(df, Diagnosis)

df %>% count(mean(Age, na.rm = TRUE), sort = TRUE)

df %>% count(Gender, sort = TRUE)

df %>% count(Ethnicity, sort = TRUE)

ggplot(df, aes(fill = Diagnosis, x = Ethnicity)) + 
  geom_bar(position = "dodge")


mean(df$Socialization, na.rm = T)
plyr::ddply(df, c("Diagnosis"), summarise, mean=mean(Socialization, na.rm = TRUE))

```

The sample included mostly young (<20) white males ...

[REPORT THE RESULTS]

## Let's test hypothesis 1: Children with ASD display a language impairment  (Exercise 2)

### Hypothesis: The child's MLU changes: i) over time, ii) according to diagnosis

Let's start with a simple mixed effects linear model

Remember to plot the data first and then to run a statistical test.
- Which variable(s) should be included as fixed factors?
- Which variable(s) should be included as random factors?

```{r ex2, include = FALSE}
library(lmerTest)

# Subsetting the columns that we want to work with:
subset <- select(df, SUBJ, VISIT, ADOS, MullenRaw, ExpressiveLangRaw, Socialization)

# Overwriting the clinical scores from visits 2-6 with the scores from visit 1 for each participant:

# have a dataset with 1 row per name from the 1st visit only and when it merges with the merged_df (after having renamed ADOS to ADOS1 etc of course) the visit 1 values should be copied to the other 5 visits as well in columns ADOS1 etc. and stay original in the original ADOS column. remember to remove column VISIT in subset_2f before merging with merged_df.

# only including visit 1 in the subset
subset <- filter(subset, VISIT == 1)

# deleting the visit column
subset <- select(subset, -VISIT)

# rename clinical measures variables
subset <- dplyr::rename(subset, ADOS2 = ADOS)
subset <- dplyr::rename(subset, MullenRaw2 = MullenRaw)
subset <- dplyr::rename(subset, ExpressiveLangRaw2 = ExpressiveLangRaw)
subset <- dplyr::rename(subset, Socialization2 = Socialization)


# merging the subset_2f with merged_df
new_df <- merge(df, subset, by = "SUBJ")

### PLOTTING

ggplot(df, aes(VISIT, CHI_MLU, colour = Diagnosis)) + 
  geom_point() + 
  geom_smooth(method = lm) +
  ggtitle('Development of MLU for children with ASD and TD over time')

### CREATING THE MODELS


 # Putting ADOS as a random slope (since we found out in portfolio 1, that ASD and TD children have roughly the same baseline lexicon size, but the TD children develop faster. Thus, the severity should have a say in the slope.). This is not possible though: you can only have within-subject factors as random slpoes.
#lm <- lmer(CHI_MLU ~ VISIT*Diagnosis + (1 + ADOS | SUBJ), df) #random slope doesnt work

lm1 <- lmer(CHI_MLU ~ VISIT*Diagnosis + (1 | SUBJ), df)
lm2 <- lmer(CHI_MLU ~ VISIT*Diagnosis + (1 + MOT_MLU | SUBJ), df)
lm3 <- lmer(CHI_MLU ~ VISIT*Diagnosis + (1 + Diagnosis + MOT_MLU | SUBJ), df)
lm4 <- lmer(CHI_MLU ~ VISIT*Diagnosis + (1 + Diagnosis | SUBJ), df)
# We decide to use these models

lm1
lm2


# Model summary

summary(lm1)
summary(lm2)

library(MuMIn)

r.squaredGLMM(lm1)
r.squaredGLMM(lm2)
r.squaredGLMM(lm3)

anova(lm1, lm2, lm4, lm3)

#We base our report of the output on the BIC values and the significance codes
#The best model is lm3, the one with two random slopes

```

How would you evaluate whether the model is a good model?

```{r ex2 evaluate, include = FALSE}

#Looking at the Rsquared. The bigger the Rsquared, the better the model is (the more variance is explained)

```

Not too good, right? Let's check whether a growth curve model is better.
Remember: a growth curve model assesses whether changes in time can be described by linear, or quadratic, or cubic (or... etc.) components.
First build the different models, then compare them to see which one is better.

```{r ex2 growth curve, include = FALSE}

df$VISIT2 <- (df$VISIT)^2
df$VISIT3 <- (df$VISIT)^3

#quadratic_m1 <- lmer(CHI_MLU ~ VISIT + VISIT2 + Diagnosis + VISIT*Diagnosis, df)
quadratic_m2 <- lmer(CHI_MLU ~ VISIT + VISIT2 + Diagnosis + VISIT*Diagnosis + (1 + Diagnosis + MOT_MLU | SUBJ), df)

summary(quadratic_m2)
r.squaredGLMM(quadratic_m2)

cubic_m2 <- lmer(CHI_MLU ~ VISIT + VISIT3 + Diagnosis + VISIT*Diagnosis + (1 + MOT_MLU | SUBJ), df)

summary(cubic_m2)

r.squaredGLMM(cubic_m2)
#R2c shows the variance explained by the whole model, R2m is the variance explained only by the fixed models.

#Why does the cubic, qudratic and linear model have the same R2c 0.827?

#Comparing the models using anova to see which one is better
anova(quadratic_m2, cubic_m2) #AIC is almost the same, as well as BIC. Why?????
anova(cubic_m2, lm2) #cubic_m2 is better than m2, and slightly better than quadratic_m2

#Comparing AIC and BIC directly
AIC(lm1, quadratic_m2, cubic_m2)
BIC(lm1, quadratic_m2, cubic_m2) #Now the quadratic seems better than the cubic WTF

```

Exciting right? Let's check whether the model is doing an alright job at fitting the data. Plot the actual CHI_MLU data against the predictions of the model fitted(model). 

```{r}

install.packages("ggplot2")
library(ggplot2)

quadratic_plot <- ggplot(df, aes(x=VISIT, y=CHI_MLU, colour=Diagnosis)) +
  geom_point()+
  geom_smooth(method = "lm", formula = y ~ poly(x, 2))+
  ggtitle("Children language development across time predicted by diagnosis - quad")

plot(quadratic_plot)

cubic_plot <- ggplot(df, aes(x=VISIT, y=CHI_MLU, colour=Diagnosis)) +
  geom_point()+
  geom_smooth(method = "lm", formula = y ~ poly(x, 3))+
  ggtitle("Children language development across time predicted by diagnosis - cubic")

plot(cubic_plot)

fitted(quadratic_m2) #should we use this function?

```

Now it's time to report our results.
Remember to report:
- the estimates for each predictor (beta estimate, standard error, p-value)
- A plain word description of the results
- A plot of your model's predictions (and some comments on whether the predictions are sensible)

[REPORT THE RESULTS]
Linguistic development of children MLU is affected by ... [COMPLETE]

```{r}
#Linguistic development of children MLU is affected by:
#  -their diagnosis (if they are diagnosed with ASD they are more prone to develop slower than TD children in regard to MLU)
#  -time (the more time passes, the more they develop in regard to MLU)
#  -their individual differences (random effects: SUBJ)

#To make sure we don't violate any assumption, we will check one by one:

#  -Normally distributed data
#  -

plot(quadratic_m2)

# Looking at plot we can see that the  assumption of linearity seems to be met since the plot line is close to zero, and the scattering of the dots suggests homoskedasticity. The mean value of VIF (for the linear model though) is 2.06 which suggests little to moderate multicollinearity. ?????????????????

```


## Let's test hypothesis 2: Parents speak equally to children with ASD and TD  (Exercise 3)

### Hypothesis: Parental MLU changes: i) over time, ii) according to diagnosis

```{r ex3, include = FALSE}

model1 <- lmer(MOT_MLU ~ CHI_MLU + VISIT*Diagnosis + (1| SUBJ), df)
model2 <- lmer(MOT_MLU ~ VISIT + CHI_MLU*Diagnosis + (1|SUBJ), df)
model3 <- lmer(MOT_MLU ~ CHI_MLU + VISIT*Diagnosis + (1+Diagnosis|SUBJ), df)
model4 <- lmer(MOT_MLU ~ CHI_MLU + VISIT*Diagnosis + (1 + Diagnosis + CHI_MLU | SUBJ), df)
model5 <- lmer(MOT_MLU ~ VISIT*Diagnosis + (1 + Diagnosis + CHI_MLU | SUBJ), df)
  
r.squaredGLMM(model1) 
r.squaredGLMM(model2)
r.squaredGLMM(model3) 
r.squaredGLMM(model4) #best Rsquared
r.squaredGLMM(model5) #2nd best Rsquared

ggplot(df, aes(VISIT, MOT_MLU))

```

Parent MLU is affected by ... but probably not ...
[REPORT THE RESULTS]

### Adding new variables (Exercise 4)

Your task now is to figure out how to best describe the children linguistic trajectory. The dataset contains a bunch of additional demographic, cognitive and clinical variables (e.g.verbal and non-verbal IQ). Try them out and identify the statistical models that best describes your data (that is, the children's MLU). Describe how you selected the best model and send the code to run the model to Victor and Byurakn.


```{r ex4, include = FALSE}



```

In addition to ..., the MLU of the children is also correlated with ...
Using AIC / nested F-tests as a criterium, we compared models of increasing complexity and found that ...

[REPORT THE RESULTS]